% LaTeX template for Artifact Evaluation V20240722
%
% Prepared by Grigori Fursin with contributions from Bruce Childers,
%   Michael Heroux, Michela Taufer and other colleagues.
%
% See examples of this Artifact Appendix in
%  * ASPLOS'24 "PyTorch 2: Faster Machine Learning Through Dynamic Python Bytecode Transformation and Graph Compilation":
%      https://dl.acm.org/doi/10.1145/3620665.3640366
%  * SC'17 paper: https://dl.acm.org/citation.cfm?id=3126948
%  * CGO'17 paper: https://www.cl.cam.ac.uk/~sa614/papers/Software-Prefetching-CGO2017.pdf
%  * ACM ReQuEST-ASPLOS'18 paper: https://dl.acm.org/citation.cfm?doid=3229762.3229763
%
% (C)opyright 2014-2024 cTuning.org
%
% CC BY 4.0 license
%

\documentclass{sigplanconf}

\usepackage{hyperref}

\begin{document}

\special{papersize=8.5in,11in}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% When adding this appendix to your paper,
% please remove above part
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix
\section{Artifact Appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Abstract}

This artifact contains the HOLA synthesizable reference model (SVM), the
NutShell RISC-V processor used as the design under test, a Linux boot image,
and an automated Artifact Evaluation workflow.  It reproduces the RCache
trade-off data in Figures 11--13, the phased Linux-boot plot in Figure 14, and
the RCore Chisel line counts in Table 4.  Every hardware experiment builds and
runs a Verilator simulator.  FPGA and Palladium platforms are not used.

\subsection{Artifact check-list (meta-information)}

{\small
\begin{itemize}
  \item {\bf Program:} SVM and NutShell, written in Chisel/Scala, SystemVerilog,
        C, C++, Python, and Bash.
  \item {\bf Compilation:} Mill, Chisel, LLVM/Clang or GCC, and Verilator.
  \item {\bf Binary:} The repository includes the NutShell Linux boot image;
        all simulator binaries are built by the workflow.
  \item {\bf Data set:} NutShell \texttt{ready-to-run/linux.bin}.
  \item {\bf Run-time environment:} OpenXiangShan xs-env, x86-64 Ubuntu 24.04
        Docker.
  \item {\bf Hardware:} A 16-core x86-64 host with at least 32 GiB RAM is
        recommended; no accelerator or programmable logic is required.
  \item {\bf Execution:} Serialized emulator builds and parallel Verilator runs.
  \item {\bf Metrics:} Evict miss rate, instruction miss rate, refill bandwidth
        overhead, IPC, load/store IPC, and physical Chisel lines of code.
  \item {\bf Output:} Four text tables, one generated Figure 14 PDF plot, one
        LoC table, and raw logs.
  \item {\bf Experiments:} MICRO 2026 paper Figures 11--14 and Table 4.
  \item {\bf How much disk space required (approximately)?:} 20 GiB, including
        the Docker image, dependency caches, build products, and raw logs.
  \item {\bf How much time is needed to prepare workflow (approximately)?:}
        20--40 minutes, depending on
        network and package-cache speed.
  \item {\bf How much time is needed to complete experiments (approximately)?:}
        2--4 hours on a 16-core host.
  \item {\bf Publicly available?:} The artifact and pinned sources are on
        \href{https://github.com/poemonsense/micro26-hola-difftest-ae}{GitHub};
        an evaluation archive supports anonymous or offline access.
  \item {\bf Code licenses (if publicly available)?:} Apache License 2.0 for SVM,
        Mulan PSL v2 for NutShell and difftest, and the license files shipped
        with other nested dependencies.
  \item {\bf Workflow automation framework used?:} Fail-fast Bash scripts plus
        Python reporting.
\end{itemize}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Description}

\subsubsection{How to access}

Obtain the artifact as a Git checkout and initialize its submodules
recursively.  The evaluated revisions are:
\begin{itemize}
  \item NutShell branch \texttt{svm}:\\
        \texttt{c1394855f28b301f5b2cb835ffa642e97bec43c4}
  \item SVM branch \texttt{main}:\\
        \texttt{0c5299a55f9a9c57acc81be36467b18eb7faa96f}
\end{itemize}
Clone the public repository recursively, or enter an existing checkout and
initialize all nested sources:

\begin{verbatim}
git clone --recursive \
  https://github.com/poemonsense/\
micro26-hola-difftest-ae hola-ae
cd hola-ae
git submodule update --init \
  --recursive
\end{verbatim}

\subsubsection{Hardware dependencies}

The workflow requires an x86-64 machine capable of running Docker.  We
recommend at least 16 logical CPU cores, 32 GiB of memory, and 20 GiB of free
disk space.  The default run concurrency is four Verilator processes.  The
experiments do not use an FPGA, Palladium, VCS, or any commercial EDA tool.

\subsubsection{Software dependencies}

The only host dependency is Docker.  The OpenXiangShan xs-env image provides
the compiler toolchain, OpenJDK, Mill, and Verilator 5.048 or newer.  Inside the
container, \texttt{setup/setup.sh} verifies those tools and installs only
missing supporting packages such as Matplotlib and initializes the two source
submodules.

\subsubsection{Data sets}

All simulations boot \texttt{NutShell/ready-to-run/linux.bin}.  The workload
boots Linux through the \texttt{/init} process, prints
\texttt{Hello, RISC-V World!}, and exits through the simulator good-trap path.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Installation}

Pull the evaluated OpenXiangShan Ubuntu 24.04 image:

\begin{verbatim}
docker pull \
ghcr.io/openxiangshan/xs-env:ubuntu-24.04
\end{verbatim}

From the artifact repository root (the directory containing
\texttt{run-all.sh}), run the complete workflow in a fresh container.  The
repository is mounted so that results remain on the host after the container
exits; the command does not depend on the host directory name.

\begin{verbatim}
docker run --rm --name hola-ae \
  -v "$PWD:/ae" -w /ae \
ghcr.io/openxiangshan/xs-env:ubuntu-24.04 \
  bash -lc './run-all.sh'
\end{verbatim}

The container needs network access during setup for any missing Ubuntu
packages, Spike, and uncached Scala dependencies.  No network access is
required after
\texttt{setup/setup.sh} has completed successfully.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment workflow}

\texttt{run-all.sh} invokes the following four fail-fast stages:

\begin{sloppypar}
\begin{enumerate}
  \item \path{setup/setup.sh} validates dependencies, initializes both
        upstream projects through their \texttt{make init} targets, and checks
        all fixed revisions.
  \item \path{build/build.sh} elaborates NutShell once, then serially builds
        each unique RCache configuration.  Every temporary source tree,
        generated RTL, compiler output, emulator, and log lives below a
        configuration-specific \path{build/build-*} directory.  Successful
        configurations are fingerprinted and can be resumed safely.
  \item \path{run/run.sh} boots Linux on every emulator.  Independent runs
        execute in parallel and write raw logs below \path{results/raw}; the
        requested \path{results/figure11} through
        \path{results/figure14} directories contain links to the relevant
        raw points.  A run is accepted only after a good trap, except for the
        documented 256 KiB/4-way capacity abort in Figure 11.
  \item \path{report/report.sh} derives all metrics from cumulative counter logs,
        validates the Figure 14 time series, invokes
        \path{SVM/scripts/plot.py}, and counts Table 4 directly from the Scala
        source.
\end{enumerate}
\end{sloppypar}

The stages may also be run separately:

\begin{verbatim}
bash setup/setup.sh
bash build/build.sh
bash run/run.sh
bash report/report.sh
\end{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation and expected results}

A successful complete run creates the following files:

\begin{verbatim}
evaluation/figure11.txt
evaluation/figure12.txt
evaluation/figure13.txt
evaluation/figure14.txt
evaluation/figure14.pdf
evaluation/table4.txt
\end{verbatim}

Each text report contains both measured values and the paper values, followed
by the maximum absolute difference in percentage points.  The expected paper
data are shown below.

\paragraph{Figure 11.}
Evict miss rate is
\[
100\times\mathtt{cache\_evict\_miss}/\mathtt{cache\_evict}.
\]

\begin{center}
\begin{tabular}{lrrr}
\hline
Cache & 4-way & 8-way & 16-way \\
\hline
256 KiB & X & 2.76\% & 3.42\% \\
512 KiB & 0.60\% & 0.41\% & 0.36\% \\
1 MiB   & 0.06\% & 0.01\% & 0.00\% \\
2 MiB   & 0.02\% & 0.00\% & 0.00\% \\
\hline
\end{tabular}
\end{center}

The X is an expected abort because every block becomes dirty and
non-replaceable.  The supplementary PLRU point reported in the paper text is
0.53\% for the 256 KiB/8-way configuration.

\paragraph{Figure 12.}
Instruction miss rate is
\[
100\times\mathtt{core\_out\_miss}/\mathtt{core\_out}.
\]

\begin{center}
\begin{tabular}{lrrrr}
\hline
Configuration & 256 KiB & 512 KiB & 1 MiB & 2 MiB \\
\hline
2-bank, 2-port & 18.32\% & 2.08\% & 0.20\% & 0.19\% \\
2-bank, 3-port & 18.18\% & 1.89\% & 0.02\% & 0.00\% \\
4-bank, 2-port & 18.27\% & 2.00\% & 0.12\% & 0.11\% \\
4-bank, 3-port & 18.18\% & 1.89\% & 0.01\% & 0.00\% \\
\hline
\end{tabular}
\end{center}

\paragraph{Figure 13.}
The disabled and enabled rows use the Figure 12 miss metric.  Bandwidth
overhead is
\[
100\times
\frac{\mathtt{cache\_read\_cache\_miss}}
     {\mathtt{cache\_refill}+\mathtt{cache\_evict}}.
\]

\begin{center}
\begin{tabular}{lrrrr}
\hline
Series & 256 KiB & 512 KiB & 1 MiB & 2 MiB \\
\hline
Disabled & 18.32\% & 2.08\% & 0.20\% & 0.19\% \\
Enabled  & 0.70\% & 0.26\% & 0.19\% & 0.19\% \\
Overhead & 2.70\% & 0.41\% & 0.00\% & 0.00\% \\
\hline
\end{tabular}
\end{center}

\paragraph{Figure 14.}
\path{evaluation/figure14.pdf} contains exactly the paper's four displayed
series: 2-port miss rate, 3-port miss rate, IPC, and load/store IPC.  The plot
should cover approximately $300\times2^{16}$ reference cycles and show the
large early 2-port spikes removed by the third port.  The plotting command does
not add the other diagnostic performance-counter series available in
\texttt{plot.py}.

Both Figure 14 runs use a 512 KiB, 8-way cache with refill-on-read-miss.  The
2-port point uses two banks; the 3-port point uses the paper's four-bank,
three-port design.

The evaluated build keeps the RTL's 1,024-cycle counter snapshots but disables
the bootrom's final UART counter dump.  This Figure 14 variant suppresses final
performance-counter output, avoiding a long diagnostic dump after the Linux
workload while retaining the samples required by \path{plot.py}.

\paragraph{Table 4.}
The expected module counts are 253, 61, 52, 126, 87, 154, 49, 25, and 129
lines for MMU, FETCH, AGU, LSU, ARITH, PRIV, TRAP, COMMIT, and RCore,
respectively, totaling 936 physical Chisel source lines.  The report excludes
the later Figure 14-only instrumentation and fails if any source boundary has
changed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment customization}

Set \path{AE_RUN_JOBS=N} to change simulator concurrency.  Set
\path{AE_RUN_TIMEOUT} to change the per-run timeout in seconds (default:
3,600).  A
space-separated subset of configuration identifiers may be supplied through
\path{AE_CONFIGS}; identifiers are listed in
\path{experiments/configs.tsv}.  For example:

\begin{verbatim}
AE_CONFIGS="c512k-w8-b2-p2-r1 \
 c512k-w8-b4-p3-r1" \
 bash build/build.sh
AE_CONFIGS="c512k-w8-b2-p2-r1 \
 c512k-w8-b4-p3-r1" \
 AE_RUN_JOBS=2 bash run/run.sh
bash report/report-figure14.sh
\end{verbatim}

Set \path{FORCE_REBUILD=1} or \path{FORCE_RERUN=1} to ignore successful
fingerprints.  Set \path{KEEP_BUILD_WORK=1} to retain complete temporary
source and compiler trees for debugging.  \path{FIGURE14_XMAX} changes the
plot's maximum x-axis value (default 305).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% When adding this appendix to your paper,
% please remove below part
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}
